---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mortality

<!-- badges: start -->
<!-- badges: end -->

This R-package fetches and organises the human mortality data from the [Human Mortality Database](mortality.org) in a tidy fashion. For a more richer and complete package, you should have a look at the `demography` package. 

The `mortality` package is a modern re-imagination and extension of the functions `hmd.mx`, `hmd.e0`, and `hmd.pop` in the `demography` package. Specifically:

* import of `RCurl` is replaced by `curl` where the latter is a modern version of the former with zero import,
* the fetched data is a special type of `tibble`, called `humble` (think human life is short so stay humble -- okay, I'll see my way out 😅), instead of a list of class `demogdata`, 
* it grabs data that are more than just period 1x1, and 
* it provides a helper function to convert `demogdata` to a `humble` table. 

## Installation

You can install the development version of mortality as below:

``` r
install.packages("remotes")
remotes::install_github("emitanaka/mortality")
```

## Usage

```{r}
library(mortality)
```


You first need to [register as a user](https://www.mortality.org/mp/auth.pl) (if you haven't already) at the Human Mortality Database. To now set a session with 
your username and password, fill in your details below and run the code. 

```r
hmd_session(username = <YOUR USERNAME>,
            password = <YOUR PASSWORD>)
```

Alternatively, you can store your username and password in the `.Renviron` file
containing below:

```
HMD_USERNAME=<YOUR USERNAME>
HMD_PASSWORD=<YOUR PASSWORD>
```

then simply just run the command before just once before getting the data:

```{r}
hmd_session()
```

All data are obtained via `hmd_data()`. 

```{r}
hmd_data("AUS", stats = "death")
```

The function offers support for multiple countries:

```{r}
hmd_data(c("AUS", "JPN"), stats = "death")
```

or multiple statistics:

```{r}
hmd_data(c("AUS", "JPN"), stats = c("death", "death_rate"))
```

or relabeling of the countries like below:

```{r}
hmd_data(c("Australia" = "AUS", "Japan" = "JPN"), stats = "population")
```

You can also get a different age or year range:

```{r}
hmd_data(c("Australia" = "AUS", "Japan" = "JPN"), 
         stats = "exposure_to_risk", year_range = 5, age_range = 5)
```
